---
title: "HW1-3"
author: "Mahdiazhari Austian & Maximilian Hein"
date: "5/3/2022"
output: html_document
---

# Homework 1 Number Three

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages


```{r cars}
library('xts') 
library('quantmod') 
library('dplyr')
library('plyr')
library("lmtest") ##load required library for grangertest
library("MTS")
library('zoo')
library('TTR')
library('imputeTS')
library('data.table')
library('tidyquant')
library('DT')
library('xtable')
```



```{r pressure, echo=FALSE}
mktsym<- c("^BSESN", "^BVSP","^FTSE","^GDAXI","^GSPC","^HSCE", 
 "^IBEX","^JKSE","^MXX","^N225","^TWII","^VLIC","^VIX")
marktCtry<-c("India", "Brazil","UK","Germany","USA","China-Shanghai",
             "Spain","Indonesia","Mexico","Japan","Taiwan","VLIC","VIX")
```



```{r}
data.env <- readRDS("data/WorldMarkts99_20.RDS")
markets <- ls(data.env)
```



```{r}
### loop through to get Ad.Close, compute weekly return and merge all stocks and treat as xts objects
returns <- xts()
per<- "weekly" ## "monthly", "daily"##period of sampling
for(i in seq_along(markets)) {
  sym <- markets[i]
  returns <- merge(returns, 
                  periodReturn(Ad(get(sym,envir=data.env)),period=per,type = "log"))
   #periodReturn(Op(get(sym,envir=data.env)),period=per,type = "arithmetic"))
}
```

```{r}
##Resolve NA's. Below a sloppy solution. Do something more intelligent
#returns[is.na(returns)]<-0 ##imputation of 0 ##or impute previous non 0 return
colnames(returns) <- paste(markets,".ret",sep="")
```



```{r}
##Extract the epoch to analysis
dI="2016-07-01"; dF="2018-06-30"
Retp <- returns[paste(dI,"/",dF,sep=""),]
```

```{r}
head(Retp)
```




```{r}
str(Retp)
```



```{r}
colnames(Retp)
```


```{r}
sum(is.na(Retp)) #number of NAs before imputation
```



```{r}
ggplot_na_distribution(Retp$JKSE.ret)
```



Use interpolation to fill in the NAs:

```{r}
for (i in 1:ncol(Retp)) {
 Retp[ , i] <- na_interpolation(Retp[,i])
  
}
```


```{r}
sum(is.na(Retp))
```


## Number 1 Calculate EMA and compare to std deviation

```{r}
ema.oneindex <- EMA(Retp$Brazil ,n=1,ratio=1-lambda)
```

```{r}
#calculate rolling standard deviation
sd.oneindex <- rollapply(Retp$Brazil, width=1, FUN=sd,na.rm=TRUE)
```

```{r}
sd.ewma <-rollapply(ema.oneindex, width=1, FUN=sd,na.rm=TRUE)
```




```{r}
#plot(sd.oneindex,t='l',col='gray20', ylab='', xlab ='',lty=3, main='Standard Deviation', cex.main=0.75)
lines(sd.oneindex,col='red',lwd=2)
lines(sd.ewma,col='green',lwd=2)
#legend('bottomright',legend = c('target','GP'),col=c('gray20','green'),lty=c(3,1),cex=.7)
```




replace column names for the countries
```{r}
colnames(Retp) <- marktCtry
```



```{r}
head(Retp)
```




```{r}
#create empty data table for storing results
causality_results_weekly <- data.table(country=marktCtry)
```


Test USA & India grangertest:
```{r}
caus1 = grangertest(Retp$USA,Retp$India,order=1)$'Pr(>F)'[2]
```


```{r}
caus1
```


```{r}
for (cty in marktCtry) {
  inter= c()
  for (cty1 in marktCtry) {
    if (cty == cty1) {
      inter= append(inter,0)
    } else {
      caus1 = grangertest(Retp[,c(as.character(cty), as.character(cty1))],order=1)$'Pr(>F)'[2]
      caus2 = grangertest(Retp[,c(as.character(cty), as.character(cty1))],order=2)$'Pr(>F)'[2]
      caus3 = grangertest(Retp[,c(as.character(cty), as.character(cty1))],order=3)$'Pr(>F)'[2]
      caus4 = grangertest(Retp[,c(as.character(cty), as.character(cty1))],order=4)$'Pr(>F)'[2]
      results = c(caus1,caus2,caus3,caus4)
      results = replace(results,results<=0.05,1)
      results = replace(results,results !=1,0)
      inter = append(inter, I(list(results)))
      
    }
  } 
 causality_results_weekly[,as.character(cty) :=inter]
  
}
```

Below is the causality result for weekly:
```{r}
causality_results_weekly
```


### Calculate volatility weekly

```{r}
#create empty data table for storing results
causality_results_weekly_volatility <- data.table(country=marktCtry)
```


```{r}
lambda=0.94
```


```{r}
weekly.volatility <- Retp
```




```{r}
for (i in 1:ncol(weekly.volatility)) {
 weekly.volatility[ , i] <- EMA(weekly.volatility[,i],n=1,ratio=1-lambda)
  
}
```




```{r}
for (cty in marktCtry) {
  inter= c()
  for (cty1 in marktCtry) {
    if (cty == cty1) {
      inter= append(inter,0)
    } else {
      caus1 = grangertest(weekly.volatility[,c(as.character(cty), as.character(cty1))],order=1)$'Pr(>F)'[2]
      caus2 = grangertest(weekly.volatility[,c(as.character(cty), as.character(cty1))],order=2)$'Pr(>F)'[2]
      caus3 = grangertest(weekly.volatility[,c(as.character(cty), as.character(cty1))],order=3)$'Pr(>F)'[2]
      caus4 = grangertest(weekly.volatility[,c(as.character(cty), as.character(cty1))],order=4)$'Pr(>F)'[2]
      results = c(caus1,caus2,caus3,caus4)
      results = replace(results,results<=0.05,1)
      results = replace(results,results !=1,0)
      inter = append(inter, I(list(results)))
      
    }
  } 
 causality_results_weekly_volatility[,as.character(cty) :=inter]
  
}
```


```{r}
 causality_results_weekly_volatility
```



## Monthly

```{r}
returns_monthly <- xts()
per<- "monthly" ## "monthly", "daily"##period of sampling
for(i in seq_along(markets)) {
  sym <- markets[i]
  returns_monthly <- merge(returns_monthly, 
                  periodReturn(Ad(get(sym,envir=data.env)),period=per,type = "log"))
   #periodReturn(Op(get(sym,envir=data.env)),period=per,type = "arithmetic"))
}
```


```{r}
colnames(returns_monthly) <- paste(markets,".ret",sep="")
```


```{r}
head(returns_monthly)
```

```{r}
##Extract the epoch to analysis
dI="2016-07-01"; dF="2018-06-30"
returns_monthly_epoch <- returns_monthly[paste(dI,"/",dF,sep=""),]
```


```{r}
sum(is.na(returns_monthly_epoch))
```


```{r}
ggplot_na_distribution(returns_monthly_epoch$JKSE.ret)
```


Use interpolation to fill in the NAs:

```{r}
for (i in 1:ncol(returns_monthly_epoch)) {
 returns_monthly_epoch[ , i] <- na_interpolation(returns_monthly_epoch[,i])
  
}
```


sum of NAs after interpolation
```{r}
sum(is.na(returns_monthly_epoch))
```


replace column names for the countries:
```{r}
colnames(returns_monthly_epoch) <- marktCtry
```


```{r}
#create empty data table for storing results
causality_results_monthly <- data.table(country=marktCtry)
```


Test USA & India grangertest:
```{r}
caus1_month = grangertest(returns_monthly_epoch$USA, returns_monthly_epoch$India,order=1)$'Pr(>F)'[2]
```


```{r}
caus1_month
```

```{r}
for (cty in marktCtry) {
  inter= c()
  for (cty1 in marktCtry) {
    if (cty == cty1) {
      inter= append(inter,0)
    } else {
      caus1 = grangertest(returns_monthly_epoch[,c(as.character(cty), as.character(cty1))],order=1)$'Pr(>F)'[2]
      caus2 = grangertest(returns_monthly_epoch[,c(as.character(cty), as.character(cty1))],order=2)$'Pr(>F)'[2]
      caus3 = grangertest(returns_monthly_epoch[,c(as.character(cty), as.character(cty1))],order=3)$'Pr(>F)'[2]
      caus4 = grangertest(returns_monthly_epoch[,c(as.character(cty), as.character(cty1))],order=4)$'Pr(>F)'[2]
      results = c(caus1,caus2,caus3,caus4)
      results = replace(results,results<=0.05,1)
      results = replace(results,results !=1,0)
      inter = append(inter, I(list(results)))
      
    }
  } 
 causality_results_monthly[,as.character(cty) :=inter]
  
}
```


```{r}
causality_results_monthly
```

### Calculate volatility monthly


```{r}
#create empty data table for storing results
causality_results_monthly_volatility <- data.table(country=marktCtry)
```


```{r}
lambda=0.94
```


```{r}
monthly.volatility <- returns_monthly_epoch
```


```{r}
for (i in 1:ncol(monthly.volatility)) {
 monthly.volatility[ , i] <- EMA(monthly.volatility[,i],n=1,ratio=1-lambda)
  
}
```


```{r}
for (cty in marktCtry) {
  inter= c()
  for (cty1 in marktCtry) {
    if (cty == cty1) {
      inter= append(inter,0)
    } else {
      caus1 = grangertest(monthly.volatility[,c(as.character(cty), as.character(cty1))],order=1)$'Pr(>F)'[2]
      caus2 = grangertest(monthly.volatility[,c(as.character(cty), as.character(cty1))],order=2)$'Pr(>F)'[2]
      caus3 = grangertest(monthly.volatility[,c(as.character(cty), as.character(cty1))],order=3)$'Pr(>F)'[2]
      caus4 = grangertest(monthly.volatility[,c(as.character(cty), as.character(cty1))],order=4)$'Pr(>F)'[2]
      results = c(caus1,caus2,caus3,caus4)
      results = replace(results,results<=0.05,1)
      results = replace(results,results !=1,0)
      inter = append(inter, I(list(results)))
      
    }
  } 
 causality_results_monthly_volatility[,as.character(cty) :=inter]
  
}
```



## Results for Monthly:
```{r}
datatable(causality_results_monthly, class = 'cell-border stripe')
```

### Monthly Volatility
```{r}
datatable(causality_results_monthly_volatility, class = 'cell-border stripe')
```



```{r}
weeklyxtable <- xtable(causality_results_weekly, auto = TRUE)
```

```{r}
weeklyxtable.volatility <- xtable(causality_results_weekly_volatility, auto = TRUE)
```

```{r}
print.xtable(weeklyxtable.volatility, type='latex', file='weekly_grangertest_volatility.tex')
```



```{r}
print.xtable(weeklyxtable, type='latex', file='weekly_grangertest.tex')

```


##Results for Weekly:
```{r}
datatable(causality_results_weekly, class = 'cell-border stripe')
```


```{r}
datatable(causality_results_weekly_volatility, class = 'cell-border stripe')
```




```{r}
monthlyxtable <- xtable(causality_results_monthly, auto = TRUE)
```


```{r}
monthlyxtable.volatility <- xtable(causality_results_monthly_volatility, auto = TRUE)
```



```{r}
print.xtable(monthlyxtable, type='latex', file='monthly_grangertest.tex')
```

```{r}
print.xtable(monthlyxtable.volatility, type='latex', file='monthly_grangertest_volatility.tex')
```

